Rate Transfers with Polyhedral Surrogate Losses

### 2021-05-12 rates-neurips-21/rates.tex
some rates work, more soon! Intro/setting

\paragraph{Target problems.}
We consider supervised machine learning problems where data points have features in $\X$ and labels in $\Y$, a finite space.
The set of probability distributions on $\Y$ is denoted $\simplex\subseteq\reals^{\Y}_+$, represented as vectors $p$ with $\|p\|_1 = 1$.

The goal or \emph{target problem} is to learn a hypothesis mapping $\X$ to a finite \emph{report (or prediction) space} $\R$, possibly different from $\Y$.
For example, in classification with an abstain option, $\R = \Y \cup \{\bot\}$ where $\bot$ represents abstaining from choosing a label.
In a ranking problem, $\R$ is the set of all permutations of $\Y$.

We suppose the target problem is specified by a \emph{discrete loss function} $\ell:\R\to\reals^\Y_+$, which maps a report (prediction) $r \in \R$ to the vector $\ell(r)$ of nonnegative loss values for each label, i.e. $\ell(r) = (\ell(r)_y)_{y\in\Y}$.
We will assume throughout that the target loss is \emph{non-redundant}, meaning every report $r \in \R$ is uniquely optimal (uniquely minimizes expected loss) for some distribution $p\in\simplex$.

\paragraph{Surrogate losses.}
Because optimizing discrete loss on a dataset is typically intractable, the usual approach is to instead learn a hypothesis of the form $h: \X \to \reals^d$ by minimizing a \emph{surrogate loss function} $L:\reals^d\to\reals^\Y_+$.
A surrogate loss is associated with a \emph{link function} $\psi: \reals^d \to \R$.
If the learned hypothesis predicts $h(x) \in \reals^d$, then the implied prediction for the target problem is $\psi(h(x)) \in \R$.
For example, in binary classification, the target problem is to predict a label from $\R = \Y = \{-1,+1\}$ and one often uses the link $\psi(u) = \mathrm{sign}(u)$, which maps negative predictions to $-1$ and nonnegative predictions to $+1$.
We will typically abstract away from the hypothesis $h$, as this paper focuses on properties of the losses $\ell(r)$ and $L(u)$ as functions of the predictions $r \in \R$ and $u \in \reals^d$.

\paragraph{Expected loss and Bayes risk.}
When $Y \sim p$, we write the expected target loss of prediction $r$ and expected surrogate loss of prediction $u$ respectively as $\inprod{p}{\ell(r)}$ and $\inprod{p}{L(u)}$.
The \emph{Bayes risk} of a discrete target loss $\ell: \R \to \reals^{\Y}_+$ is the function $\risk{\ell}: \simplex \to \reals_+$ given by $\risk{\ell}(p) = \inf_{r \in \R} \inprod{p}{\ell(r)}$.
Similarly, the Bayes risk of a surrogate $L: \reals^d \to \reals^{\Y}_+$ is given by $\risk{L}(p) = \inf_{u\in\reals^d} \inprod{p}{L(u)}$.
Note that we define Bayes risk pointwise, i.e. we have abstracted away from the role of the features $\X$ and focus directly on conditional distributions $p$ on $\Y$.

@@

In supervised machine learning we are given a data set $S_m = ((x_1,y_1), \dots, (x_m,y_m)) \sim \D^m$, that is, $m$ points drawn i.i.d. from a distribution $\D$ on $\X \times \Y$.
Write $\D_{\X}$ for the marginal distribution of $\D$ on $\X$, and for $(X,Y) \sim \D$, write $p_x$ for the conditional distribution on $Y$ given $X=x$.

Given a target loss $\ell: \R \to \reals^{\Y}_+$, our goal is to output a hypothesis $g: \X \to \R$ minimizing \emph{target risk} $E_{X \sim \D_\X} \inprod{p_X}{\ell(g(X))}$.
Then we can define the \emph{excess risk} or \emph{regret} of $g$ as the risk of $g$, minus the risk of the Bayes optimal predictor.
\begin{definition}[Regret] \label{def:full-regret}
  For a loss $\ell: \R \to \reals^{\Y}_+$ and distribution $p \in \simplex$, the \emph{conditional regret} of a prediction $r \in \R$ is given by $R_{\ell}(r,p) = \inprod{p}{\ell(r)} - \risk{\ell}(p)$.

  In slight abuse of notation, the \emph{regret} of a hypothesis $g: \X \to \R$ is given by $R_{\ell}(g, \D) = \E_{X \sim \D_{\X}} R_{\ell}(g(X), p_X)$.
\end{definition}
Naturally, the same definitions hold with the notation of surrogate losses, i.e. we have conditional regret $R_L(u,p) = \inprod{p}{L(u)} - \risk{L}(p)$ and we have regret $R_L(h,\D) = E_{X \sim \D_{\X}} R_L(h(X), p_X)$.

%The primary focus of this paper is on surrogate regret bounds, which bound the excess target loss in terms of the excess surrogate loss.
%
%\begin{definition}[Regret]\label{def:regret}
%  For loss $L:\R\to\reals^\Y_+$ and distribution $p\in\simplex$, the \emph{regret} of a prediction $r\in\R$ is given by $R_L(r,p) = \inprod{p}{L(r)} - \risk{L}(p)$.
%\end{definition}
%
%\begin{definition}[Surrogate Regret Bound]\label{def:regret-bound}
%  A \emph{surrogate regret bound} between target $\ell:\R\to\reals^\Y$ and surrogate $L:\reals^d\to\reals^\Y$, together with a link $\psi : \reals^d \to \R$, is the statement
%  \begin{equation}
%    \label{eq:regret-bound-general}
%    \forall p\in\simplex, u\in\reals^d, \quad R_\ell(\psi(u),p) \leq \zeta(\, R_L(u,p) \,)~,
%  \end{equation}
%  for some function $\zeta : \reals_+ \to \reals_+$ which is continuous at $0$ and satisfies $\zeta(0) = 0$.
%\end{definition}
%
%
%
%\subsection{Property elicitation and consistency}
%
%Given a target problem specified by $\ell$, when is it appropriate to use a surrogate $L$?
%A key necessary condition is \emph{consistency} \bo{cite}, which roughly states that 
%
%

%$\Gamma: \simplex \toto \R$ is shorthand for $\Gamma: \simplex \to 2^{\R} \setminus \{\emptyset\}$.
%
%\begin{definition}[Property, level set]\label{def:property}
%  A \emph{property} is a function $\Gamma:\simplex\toto\R$.
%  The \emph{level set} of $\Gamma$ for report $r$ is the set $\Gamma_r \defeq \{p \in \simplex : r \in \Gamma(p)\}$.
%\end{definition}
%
%\begin{definition}[Elicits]
%  \label{def:elicits}
%  A loss $L:\R\to\reals^\Y_+$, \emph{elicits} a property $\Gamma:\simplex \toto \R$ if
%  \begin{equation}
%    \forall p\in\simplex,\;\;\;\Gamma(p) = \argmin_{r \in \R} \inprod{p}{L(r)}~.
%  \end{equation}
%  As $L$ elicits a unique property, we write $\prop{L}$ to refer to the property elicited by a loss $L$.
%\end{definition}
%
%\begin{definition}[Embedding a loss]\label{def:loss-embed}
%  A loss $L:\reals^d\to\reals^\Y$ \emph{embeds} a loss $\ell:\R\to\reals^\Y$ if there exists some injective embedding $\varphi:\R\to\reals^d$ such that
%  (i) for all $r\in\R$ we have $L(\varphi(r)) = \ell(r)$, and (ii) for all $p\in\simplex,r\in\R$ we have
%  \begin{equation}\label{eq:embed-loss}
%    r \in \prop{\ell}(p) \iff \varphi(r) \in \prop{L}(p)~.
%  \end{equation}
%\end{definition}
%%
%

>>>>

\paragraph{Target problems.}
We consider supervised machine learning problems where data points have features in $\X$ and labels in $\Y$, a finite space.
The \emph{target problem} is to learn a hypothesis mapping $\X$ to a finite \emph{report (or prediction) space} $\R$, possibly different from $\Y$.
For example, in classification with an abstain option, $\R = \Y \cup \{\bot\}$ where $\bot$ represents abstaining from choosing a label.
In a ranking problem, $\R$ is the set of all permutations of $\Y$.

We suppose the target problem is specified by the \emph{target loss} $\ell:\R\to\reals^\Y_+$, which maps a report (prediction) $r \in \R$ to the vector $\ell(r)$ of nonnegative loss values for each label, i.e. $\ell(r) = (\ell(r)_y)_{y\in\Y}$.
Given a hypothesis $g: \X \to \R$ and a data point $(x,y) \in \X \times \Y$, the loss is $\ell(g(x))_y$.

\paragraph{Surrogate problems.}
Because optimizing a discrete target loss on a dataset is typically intractable, the usual approach is \emph{surrogate risk minimization}.
First, one constructs a \emph{surrogate loss function} $L:\reals^d\to\reals^\Y_+$.
Then, one minimizes it to learn a hypothesis of the form $h: \X \to \reals^d$.
Finally, one applies a \emph{link function} $\psi: \reals^d \to \R$ that is associated with the surrogate loss.
If the learned hypothesis predicts $h(x) \in \reals^d$, then the implied prediction for the target problem is $\psi(h(x)) \in \R$.

For example, in binary classification, the target problem is to predict a label from $\R = \Y = \{-1,+1\}$ with $0$-$1$ loss.
The surrogate prediction space is typically $\reals$ and common surrogate losses include hinge loss and logistic loss.
These surrogates are typically associated with the link $\psi(u) = \mathrm{sign}(u)$, which maps negative predictions to $-1$ and nonnegative predictions to $+1$.
In general, given a target problem, the question is how to design a ``good'' surrogate loss $L$ and link $\psi$.
We will address one aspect of this question in this paper.

\paragraph{Bayes risk, regret, and the conditional perspective.}
In supervised machine learning, we are given a data set $S_m = ((x_1,y_1), \dots, (x_m,y_m)) \sim \D^m$, that is, $m$ points drawn i.i.d. from a distribution $\D$ on $\X \times \Y$.
The set of probability distributions on $\Y$ is denoted $\simplex\subseteq\reals^{\Y}_+$, represented as vectors $p$ with $\|p\|_1 = 1$.
Write $\D_{\X}$ for the marginal distribution of $\D$ on $\X$, and for $(X,Y) \sim \D$, write $p_x \in \simplex$ for the conditional distribution on $Y$ given $X=x$.

We note here that, to examine properties of loss functions $\ell(r)$ and $L(u)$, we will often abstract away from the role of the features $\X$ and the hypothesis $h$.
That is, rather than considering $x$, $p_x$, $h(x)$, $L(h(x))$, and so on, we will often simply consider some conditional distribution $p \in \simplex$, some report (prediction) $u \in \reals^d$, its loss $L(u)$, and so on.

When $Y \sim p$, we write the expected target loss of prediction $r$ as $\inprod{p}{\ell(r)}$ and the expected surrogate loss of prediction $u$ as $\inprod{p}{L(u)}$.
The (conditional) \emph{Bayes risk} of a discrete target loss $\ell: \R \to \reals^{\Y}_+$ is the function $\risk{\ell}: \simplex \to \reals_+$ given by $\risk{\ell}(p) = \inf_{r \in \R} \inprod{p}{\ell(r)}$.
Similarly, the Bayes risk of a surrogate $L: \reals^d \to \reals^{\Y}_+$ is given by $\risk{L}(p) = \inf_{u\in\reals^d} \inprod{p}{L(u)}$.

Given $\ell$, the target problem is to output a hypothesis $g: \X \to \R$ minimizing \emph{target risk} $\E_{X \sim \D_\X} \inprod{p_X}{\ell(g(X))}$.
We define the \emph{excess risk} or \emph{regret} of $g$ as the risk of $g$, minus the risk of the Bayes optimal predictor:
\begin{definition}[Regret] \label{def:full-regret}
  For a loss $\ell: \R \to \reals^{\Y}_+$ and distribution $p \in \simplex$, the \emph{conditional regret} of a prediction $r \in \R$ is given by $R_{\ell}(r,p) = \inprod{p}{\ell(r)} - \risk{\ell}(p)$. \bo{Or just call it regret?}

  In slight abuse of notation, the \emph{regret} of a hypothesis $g: \X \to \R$ is written $R_{\ell}(g; \D) = \E_{X \sim \D_{\X}} R_{\ell}(g(X), p_X)$.
\end{definition}
Naturally, the same definitions hold for the surrogate problem of minimizing \emph{surrogate risk} $\E_{X \sim \D_\X} \inprod{p_X}{L(h(x))}$, i.e. we have conditional regret $R_L(u,p) = \inprod{p}{L(u)} - \risk{L}(p)$ and we have regret $R_L(h;\D) = E_{X \sim \D_{\X}} R_L(h(X), p_X)$.

\bo{We are assuming Bayes in class here. Should comment about it, maybe discuss!}

\paragraph{Surrogate transfer bounds.}
Central to supervised machine learning is the \emph{rate of convergence} of regret as a function of $m$, the number of data points.
If a learning algorithm's output is denoted $\hat{h}_m$ on $m$ data points, then a typical bound is of the form $R_L(\hat{h}_m;\D) \leq O\left(\frac{1}{m^{\alpha}}\right)$ with high probability.
If $\alpha = \frac{1}{2}$, then regret tends to zero with the inverse square root of the dataset size; this is considered a ``slow rate'' and is guaranteed e.g. by classical PAC learning bounds..
On the other hand, $\alpha = 1$ is considered a ``fast rate'' and is achieveable under a variety of conditions \bo{cite}.

However, much less is known about the relationship between surrogate and target regret.
Given a convergence rate for surrogate regret $R_L(\hat{h}_m;\D)$, what is the implied convergence rate for target regret $R_{\ell}(\hat{g}_m;\D)$, where $\hat{g}_m(x) = \psi(\hat{h}_m(x))$?
\bo{Prior work to mention here?}

To systematize the investigation of this problem, we formalize the notion of \emph{surrogate transfer bounds}.

%\begin{definition}[Surrogate Regret Bound]\label{def:regret-bound}
%  A \emph{surrogate regret bound} between target $\ell:\R\to\reals^\Y$ and surrogate $L:\reals^d\to\reals^\Y$, together with a link $\psi : \reals^d \to \R$, is the statement
%  \begin{equation}
%    \label{eq:regret-bound-general}
%    \forall p\in\simplex, u\in\reals^d, \quad R_\ell(\psi(u),p) \leq \zeta(\, R_L(u,p) \,)~,
%  \end{equation}
%  for some function $\zeta : \reals_+ \to \reals_+$ which is continuous at $0$ and satisfies $\zeta(0) = 0$.
%\end{definition}
%

@@

%The primary focus of this paper is on surrogate regret bounds, which bound the excess target loss in terms of the excess surrogate loss.
%
%\begin{definition}[Regret]\label{def:regret}
%  For loss $L:\R\to\reals^\Y_+$ and distribution $p\in\simplex$, the \emph{regret} of a prediction $r\in\R$ is given by $R_L(r,p) = \inprod{p}{L(r)} - \risk{L}(p)$.
%\end{definition}
%

%
%
%\subsection{Property elicitation and consistency}
%
%Given a target problem specified by $\ell$, when is it appropriate to use a surrogate $L$?
%A key necessary condition is \emph{consistency} \bo{cite}, which roughly states that 
%
%
%We will assume throughout that the target loss is \emph{non-redundant}, meaning every report $r \in \R$ is uniquely optimal (uniquely minimizes expected loss) for some distribution $p\in\simplex$.
%$\Gamma: \simplex \toto \R$ is shorthand for $\Gamma: \simplex \to 2^{\R} \setminus \{\emptyset\}$.
%
%\begin{definition}[Property, level set]\label{def:property}
%  A \emph{property} is a function $\Gamma:\simplex\toto\R$.
%  The \emph{level set} of $\Gamma$ for report $r$ is the set $\Gamma_r \defeq \{p \in \simplex : r \in \Gamma(p)\}$.
%\end{definition}
%
%\begin{definition}[Elicits]
%  \label{def:elicits}
%  A loss $L:\R\to\reals^\Y_+$, \emph{elicits} a property $\Gamma:\simplex \toto \R$ if
%  \begin{equation}
%    \forall p\in\simplex,\;\;\;\Gamma(p) = \argmin_{r \in \R} \inprod{p}{L(r)}~.
%  \end{equation}
%  As $L$ elicits a unique property, we write $\prop{L}$ to refer to the property elicited by a loss $L$.
%\end{definition}
%
%\begin{definition}[Embedding a loss]\label{def:loss-embed}
%  A loss $L:\reals^d\to\reals^\Y$ \emph{embeds} a loss $\ell:\R\to\reals^\Y$ if there exists some injective embedding $\varphi:\R\to\reals^d$ such that
%  (i) for all $r\in\R$ we have $L(\varphi(r)) = \ell(r)$, and (ii) for all $p\in\simplex,r\in\R$ we have
%  \begin{equation}\label{eq:embed-loss}
%    r \in \prop{\ell}(p) \iff \varphi(r) \in \prop{L}(p)~.
%  \end{equation}
%\end{definition}
%%
%
